1.This exercise relates to the College data set, which can be found in the file College.csv. It contains a number of variables for 777 different universities and colleges in the US. The variables are

Private: Public/private indicator
Apps: Number of applications received
Accept: Number of applicants accepted
Enroll: Number of new students enrolled
Top10perc: New students from top 10% of high school class
Top25perc: New students from top 25% of high school class
F.Undergrad: Number of full-time undergraduates
P.Undergrad: Number of part-time undergraduates
Outstate: Out-of-state tuition
Room.Board: Room and board costs
Books: Estimated book costs
Personal: Estimated personal spending
PhD: Percent of faculty with Ph.D.'s
Terminal: Percent of faculty with terminal degree
S.F.Ratio: Student/faculty ratio
perc.alumni: Percent of alumni who donate
Expend: Instructional expenditure per student
Grad.Rate: Graduation rate






Part A

```{r}
college = read.csv('/Users/anjan/Downloads/College.csv', header = TRUE)
```

```{r}
head(college)
```


Part 2
```{r}
rownames(college) = college[, 1]
head(college)
```


```{r}
college = college[, -1]
head(college)
```

part 3.1
```{r}
summary(college)
```


Part 3.2

```{r}
pairs(college[, 1:10])
```


Part 3.3
```{r}
plot(college$Private, college$Outstate, xlab = "Private", ylab = "Out-of-state tuition (dollars)")
```



part 3.4
```{r}
Elite = rep("No", nrow(college))
Elite[college$Top10per>50] = "Yes"
Elite = as.factor(Elite)
college = data.frame(college, Elite)
```

```{r}
Elite = rep("No", nrow(college))
Elite[college$Top10per > 50] = "Yes"
Elite = as.factor(Elite)
college = data.frame(college, Elite)
```

```{r}
summary(college$Elite)
```


```{r}
plot(college$Elite, college$Outstate, xlab = "Elite", ylab = "Out-of-state tuition (dollars)")
```
Part 3.5

```{r}
par(mfrow=c(2,2))
hist(college$Apps, xlab = "Applications Received", main = "")
hist(college$perc.alumni, col=2, xlab = "% of alumni who donate", main = "")
hist(college$S.F.Ratio, col=3, breaks=10, xlab = "Student/faculty ratio", main = "")
hist(college$Expend, breaks=100, xlab = "Instructional expenditure per student", main = "")
```
```{r}
par(mfrow=c(2,2))
hist(college$Expend, xlab = "Applications Received", main = "")
hist(college$perc.alumni, col=2, xlab = "% of alumni who donate", main = "")
hist(college$S.F.Ratio, col=3, breaks=10, xlab = "Student/faculty ratio", main = "")
hist(college$Expend, breaks=100, xlab = "Instructional expenditure per student", main = "")
```
```{r}
# Some interesting observations:
# what is the university with the most students in the top 10% of class
row.names(college)[which.max(college$Top10perc)]  
acceptance_rate <- college$Accept / college$Apps
# what university has the smallest acceptance rate
row.names(college)[which.min(acceptance_rate)]  
# what university has the most liberal acceptance rate
row.names(college)[which.max(acceptance_rate)]
# High tuition correlates to high graduation rate
plot(college$Outstate, college$Grad.Rate) 
# Colleges with low acceptance rate tend to have low S:F ratio.
plot(college$Accept / college$Apps, college$S.F.Ratio) 
# Colleges with the most students from top 10% perc don't necessarily have
# the highest graduation rate. Also, rate > 100 is erroneous!
plot(college$Top10perc, college$Grad.Rate)
```



2.

This exercise involves the Auto data set studied in the lab. Make sure that the missing values have been removed from the data.


```{r}
Auto = read.csv("/Users/anjan/Downloads/Auto.csv", header = TRUE, na.strings = "?")
Auto = na.omit(Auto)
dim(Auto)
```

Part 1


```{r}
head(Auto)
```

Part 2
```{r}
range(Auto$mpg)

```

```{r}
range(Auto$cylinders)
```

```{r}
range(Auto$displacement)
```

```{r}
range(Auto$horsepower)
```

```{r}
range(Auto$weight)
range(Auto$acceleration)
range(Auto$year)
```

Part 3 

```{r}
colMeans(Auto[, 1:7])
```


Part 4

```{r}
apply(Auto[, 1:7], MARGIN = 2, FUN = "sd")
```
```{r}
apply(Auto[-(10:85), 1:7], MARGIN = 2, FUN = "range")
```
```{r}
apply(Auto[-(10:85), 1:7], MARGIN = 2, FUN = "mean")

```

```{r}
apply(Auto[-(10:85), 1:7], MARGIN = 2, FUN = "sd")

```

Part 5

```{r}
par(mfrow = c(2, 2))
plot(Auto$displacement, Auto$mpg, xlab = "Engine displacement (cubic inches)", ylab = "Miles per gallon")
plot(Auto$horsepower, Auto$mpg, xlab = "Horsepower", ylab = "Miles per gallon")
plot(Auto$weight, Auto$mpg, xlab = "Car weight (pounds)", ylab = "Miles per gallon")
plot(Auto$year, Auto$mpg, xlab = "Model Year", ylab = "Miles per gallon")
```
```{r}
par(mfrow = c(2, 2))
plot(Auto$year, Auto$acceleration, xlab = "Model Year", ylab = "0 to 60mph time (seconds)")
plot(Auto$year, Auto$displacement, xlab = "Model Year", ylab = "Engine displacement (cubic inches)")
plot(Auto$year, Auto$weight, xlab = "Model Year", ylab = "Car weight (pounds)")
plot(Auto$year, Auto$horsepower, xlab = "Model Year", ylab = "Horsepower")
```

```{r}
par(mfrow = c(2, 2))
plot(Auto$weight, Auto$acceleration, xlab = "Car weight (pounds)", ylab = "0 to 60mph time (seconds)")
plot(Auto$cylinders, Auto$acceleration, xlab = "Number of engine cylinders", ylab = "0 to 60mph time (seconds)")
plot(Auto$displacement, Auto$acceleration, xlab = "Engine displacement (cubic inches)", ylab = "0 to 60mph time (seconds)")
plot(Auto$horsepower, Auto$acceleration, xlab = "Horsepower", ylab = "0 to 60mph time (seconds)")
```
```{r}
par(mfrow = c(2, 1))
plot(Auto$weight, Auto$horsepower, xlab = "Car weight (pounds)", ylab = "Horsepower")
plot(Auto$weight, Auto$displacement, xlab = "Car weight (pounds)", ylab = "Engine displacement (cubic inches)")
```
Part 6

```{r}
Auto$origin[Auto$origin == 1] = "American"
Auto$origin[Auto$origin == 2] = "European"
Auto$origin[Auto$origin == 3] = "Japanese"
Auto$origin = as.factor(Auto$origin)
plot(Auto$origin, Auto$mpg, xlab = "Country of origin", ylab = "Miles per gallon")
```

3.This exercise involves the Boston housing data set.
```{r}
library(MASS)
```
Part 1

```{r}
head(Boston)

```
```{r}
?Boston

```

Part 2 

```{r}
dim(Boston)

```
```{r}
attach(Boston)

pairs(Boston)
```
```{r}
Boston.corr = cor(Boston)
Boston.corr.crim = Boston.corr[-1,1]
print(
  Boston.corr.crim[order(abs(Boston.corr.crim), decreasing = T)]
)
```
Part 3 
```{r}
par(mfrow=c(2,2))
# get the four most correlated variables
aux = names(Boston.corr.crim[order(abs(Boston.corr.crim), decreasing = T)][1:4])
for(i in aux){
  plot(get(i), crim, xlab=i)
}
```
Part 4
```{r}
summary(crim)
```

```{r}
length(crim[crim>30])
```

```{r}
hist(tax)
```
```{r}
length(tax[tax>500])
```

```{r}
hist(ptratio)
```
```{r}
length(ptratio[ptratio<14])
```

```{r}
table(chas)
```

```{r}
median(ptratio)
```

```{r}
subs.lw = which(medv<median(medv))
print(subs.lw)
```

```{r}
Boston.corr.subs.lw = cor(Boston[subs.lw, ])
corr.compare = data.frame('lower'=Boston.corr.subs.lw[, "medv"], 'all'=Boston.corr[, "medv"])
corr.compare$diff = corr.compare$lower - corr.compare$all
```

```{r}
hist(corr.compare$diff, xlab="Correlation Differences")
```

```{r}
hist(abs(corr.compare$diff), xlab="Correlation Differences")
```

```{r}
main.diffs = head(corr.compare[order(abs(corr.compare$diff), decreasing = T), ], 5)

print(main.diffs)
```

```{r}
print(rownames(main.diffs))
```

```{r}
hist(rm, main="Distribution of Rooms by Dwelling", xlab="Rooms")
```
```{r}
length(rm[rm>7])
```

```{r}
length(rm[rm>8])
```

```{r}
frm =as.factor(as.character(lapply(rm, function(x) ifelse(x>8, "]8, +Inf[", ifelse(x>7,"]7,8]","[0,7]")))))
plot(frm, medv, varwidth=T, xlab="Number of Rooms", 
     ylab="Median Values by $1000s",
     title="Median Value of Owner-Occupied Homes")

```

```{r}
Boston[rm>8 & medv<30, ]
```

4.



This question involves the use of simple linear regression on the Auto data set.



```{r}
Auto = read.csv('/Users/anjan/Downloads/Auto.csv', header = TRUE, na.strings = "?")
Auto = na.omit(Auto)

```
```{r}
auto.lin.fit = lm(mpg ~ horsepower, data = Auto)
summary(auto.lin.fit)
```
```{r}
predict(auto.lin.fit, data.frame(horsepower = 98), interval = "confidence")
```


```{r}
predict(auto.lin.fit, data.frame(horsepower = 98), interval = "prediction")
```

```{r}
plot(Auto$horsepower, Auto$mpg, xlab = "Horsepower", ylab = "Miles per gallon")
abline(auto.lin.fit, lwd = 3, col = "red")
```
5.

This question involves the use of multiple linear regression on the Auto data set.

```{r}
par(mfrow = c(2, 2))
plot(auto.lin.fit)
```
```{r}
Auto = read.csv('/Users/anjan/Downloads/Auto.csv', header = TRUE, na.strings = "?")
Auto = na.omit(Auto)
```

```{r}
Auto$origin[Auto$origin == 1] = "American"
Auto$origin[Auto$origin == 2] = "European"
Auto$origin[Auto$origin == 3] = "Japanese"
Auto$origin = as.factor(Auto$origin)
head(Auto)
```
```{r}
pairs(~mpg + cylinders + displacement + horsepower + weight + acceleration + year, Auto)
```
```{r}
cor(Auto[,-c(8, 9)])
```

```{r}
mpg.fit = lm(mpg ~ . - name, data = Auto)
summary(mpg.fit)
```
```{r}
contrasts(Auto$origin)

```

```{r}
par(mfrow = c(2, 2))
plot(mpg.fit)

```
```{r}
mpg.fit.all.interactions = lm(mpg ~ (. - name)^2, data = Auto)
summary(mpg.fit.all.interactions)
```
```{r}
mpg.fit.reduced.interactions = update(mpg.fit.all.interactions, ~ . - horsepower:origin)
summary(mpg.fit.reduced.interactions)
```

```{r}
mpg.fit.reduced.interactions = update(mpg.fit.reduced.interactions, ~ . - displacement:horsepower)
summary(mpg.fit.reduced.interactions)
```
```{r}
mpg.fit.reduced.interactions = update(mpg.fit.reduced.interactions, ~ . - weight:acceleration)
summary(mpg.fit.reduced.interactions)
```
```{r}
mpg.fit.reduced.interactions = update(mpg.fit.reduced.interactions, ~ . - cylinders:origin)
summary(mpg.fit.reduced.interactions)

```

```{r}
mpg.fit.reduced.interactions = update(mpg.fit.reduced.interactions, ~ . - cylinders:displacement)
summary(mpg.fit.reduced.interactions)
```
```{r}
mpg.fit.reduced.interactions = update(mpg.fit.reduced.interactions, ~ . - cylinders:weight)
summary(mpg.fit.reduced.interactions)
```
```{r}
summary(lm(mpg ~ . + cylinders:displacement - name, data = Auto))

```
```{r}
par(mfrow = c(2, 2))
plot(Auto$displacement, Auto$mpg)
plot(Auto$horsepower, Auto$mpg)
plot(Auto$weight, Auto$mpg)
plot(Auto$acceleration, Auto$mpg)
```
```{r}
summary(lm(mpg ~ acceleration, data = Auto))

```
```{r}
par(mfrow = c(2, 2))
plot(lm(mpg ~ acceleration, data = Auto))
```
```{r}
summary(lm(mpg ~ log(acceleration), data = Auto))

```
```{r}
par(mfrow = c(2, 2))
plot(lm(mpg ~ log(acceleration), data = Auto))
```
```{r}
displacement.linear = lm(mpg ~ displacement, data = Auto)
summary(displacement.linear)
```

```{r}
displacement.quadratic = lm(mpg ~ poly(displacement, 2), data = Auto)
summary(displacement.quadratic)
```

```{r}
anova(displacement.linear, displacement.quadratic)

```
```{r}
displacement.quintic = lm(mpg ~ poly(displacement, 5), data = Auto)
summary(displacement.quintic)
```
```{r}
displacement.quintic = lm(mpg ~ poly(displacement, 5), data = Auto)
summary(displacement.quintic)
```
```{r}
anova(displacement.quadratic, displacement.quintic)
```
6.


This question should be answered using the Carseats data set.
```{r}
library(ISLR)
head(Carseats)
```
```{r}
carseats.fit.1 = lm(Sales ~ Price + Urban + US, data = Carseats)
summary(carseats.fit.1)
```
```{r}
contrasts(Carseats$Urban)
```

```{r}
contrasts(Carseats$US)
```














```{r}
carseats.fit.2 = lm(Sales ~ Price + US, data = Carseats)
summary(carseats.fit.2)
```
```{r}
par(mfrow = c(2, 2))
plot(carseats.fit.1)
```
```{r}
par(mfrow = c(2, 2))
plot(carseats.fit.2)
```



```{r}
confint(carseats.fit.2)
```

